---
title: "Project 3"
author: "Damon D'Ambrosio"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
---

## App Description and Purpose

The purpose of this app is to see if various data associated with Pokemon (i.e., their base stats, height, and weight), can be used to predict their type. Type membership is determined in a binary fashion. Since each Pokemon can have up to two types, a primary and secondary, the app will consider both types when assigning type membership. For example, a Pokemon with a primary type of "Fire" would be considered a member of the "Fire" type, as would a different Pokemon with a primary type of "Dark" and a secondary type of "Fire." Data originate from [PokeApi](https://pokeapi.co/). For the purpose of expediting load time for this app, the dataset resides locally in this GitHub repo.

Data can be subset as the user sees fit to limit the scope of included Pokemon by "Generation" (i.e., the series of games in which they appeared), and different predictors can be included/excluded from each model. When applicable, certain tuning parameters can be specified for the models. The app will fit a generalized linear model, a simple classification tree, and a random forest ensemble model.

These data are inherently imbalanced, as there are always more Pokemon that do NOT belong to a specific type versus those that do. As such, models often superficially appear to have high accuracy, as simply "guessing" most Pokemon do NOT belong to a certain type will artificially inflate this value. To guard against this, the app provides other fit metrics, such as "sensitivity" and "specificity," to give a better idea of model performance, and explicitly identifies a Pokemon belonging to a given type as the "positive." Because of this imbalance, sometimes, unique combinations of type, Generation filter, tuning parameters, training/test split, and selected predictors will fail to produce some models (epsecially the GLM/simple classification tree). Some safeguards exist within the app to preclude the user from trying to fit a model on a filtered dataset that has no "positive" occurences of type membership.

In addition to making models, the app allows for numeric and graphical exploration of these data, allows the user to provide custom inputs (fully user-defined, as well as lists of Pokemon that are and are not members of a selected type) to see model predictions, and allows the user to filter and export the data itself as a .csv file.

## List of required packages

The following packages are needed to run this app:

- `shiny`: Create the Shiny app itself
- `tidyverse`: Data manipulation/figure-making
- `caret`: Create/train the various models
- `randomForest`: Create the random forest model
- `DT`: Create the `dataTableOutput`-type objects within the app

The following code will install these packages:

```{r dependencies, eval= FALSE}
install.packages(
  c("shiny",
    "tidyverse",
    "caret",
    "randomForest",
    "DT"
  )
)
```

## Running the app
Assuming all dependencies are installed, the folllowing code will run the app:

```{r run, eval = FALSE}
shiny::runGitHub('project-3', 'dadambro')
```
